### 2012-10-27 intention motor representation joint action (paper 3)/conf budapest philosophy 2012-11/tmp.tex
before change introduction messing up contrast cases

\section{Shared intention}
On the leading, best developed accounts of shared agency in philosophy and psychology, 
shared agency is explained  in terms of a special kind of intention or structure of intention, knowledge and commitment often called a shared intention.
For two or more agents to exercise shared agency is for them to have a shared intention and for this shared intention to be appropriately related to their actions.
Such actions are called a `shared intentional activity' (\citealp[p.\ 142]{Bratman:1999fr}).%
%On this widely held view, the difference between our walking together and two strangers merely walking side-by-side is that one event is appropriately related to a shared intention whereas the other is not.

@@

To avoid the need to take sides on what precisely shared intention is, 
let us abstract from some details. 
There are features of shared intention on which there should be widespread agreement.
Minimally, a shared intention stands to the actions comprising a shared intentional activity in roughly the way that an ordinary, individual intention stands to an ordinary, individual action.
Now an ordinary, individual intention represents an outcome, coordinates an agent's actions, and coordinates the agent's actions in such a way that, normally, this coordination would facilitate the occurrence of the represented outcome.
Similarly, 
on almost any account,
a shared intention involves there being a single outcome which each agent represents, where these representations coordinate the several agents' actions and coordinate them in such a way that, normally, the coordination would facilitate the occurrence of the represented outcome.

Our view involves rejecting the claim that a full account of shared agency can be given in terms of shared intention alone.
Instead, we shall argue, it is also necessary to appeal to certain structures of motor representation. 
Note, however, that our view is compatible with various weaker claims linking shared intention to shared agency.
In particular, our view is consistent with the claims that appeal shared intention is also needed in giving an account of shared agency, 
and that \emph{some forms} of shared agency can be fully characterised by appeal to shared intention alone.
Our view is also consistent with the claim that all exercises of shared agency involve shared intentions in some way.%
\footnote{
See \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  
Related claims are made by 
	\citet[p.\ 381]{Carpenter:2009wq}, 
	\citet[p.\ 369]{Call:2009fk}, 
	\citet{Kutz:2000si},  
	\citet[p.\ 117]{rakoczy_pretend_2006},  and
	\citet{Tollefsen:2005vh}.
}

@@

First, if someone thought that it is possible to distinguish systematically between the contrast cases by appeal to shared intention (and perhaps only by appeal to shared intention), this could motivate holding that for an event to be a joint action is for it to be appropriately related to a shared intention.
We shall argue, however, that distinguishing systematically between the contrast cases cannot be done entirely in terms of shared intention. 
[*CV: this argument is missing from this draft.]

{\label{text:second_consideration} A second, related consideration} which might appear to favour this view is a parallel between shared agency and ordinary, individual agency.
If a full account of individual agency could be given by appeal to intention and other propositional attitudes with no appeal to motor representation,
then it would seem unlikely that giving a full account of shared agency must somehow involve appeal to motor representation.
However, as mentioned above, some have argued that fully characterising individual agency will require appeal to motor representation as well as intention \citep{pacherie:2000_content,butterfill:2012_intention}. 
Given these arguments, the parallel between individual and shared agency motivates considering the possibility that  there is a constitutive role for motor representation in shared agency. 
We also note that some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
For instance, Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' (\citeyear[p.\ 251]{bratman:2000_valuing}; see \citealp[pp.\ 137--8]{Bratman:1987xw} for further justification). 
Certainly, then, it is not unreasonable to question the claim that a full account of ordinary, individual agency can be given just in terms of intention.

\section{Motor representation}
Since we aim to argue that a fully adequate account of shared agency must appeal to motor representation, it may be helpful first to provide some background about motor representation.

@@

So much for the bare notion of motor representation.
Why suppose that motor representation might be needed to explain shared agency?

@@

\section{Social motor representation}
We start from an empirical premise which is about enabling conditions for shared agency only: some joint actions are facilitated by reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent.

A representation (motor or not) is \emph{agent-neutral} if its content does not specify any particular agent or agents.%
\footnote{
Our use of the term `agent-netural' to describe motor representations bears no relation to the use of the same term  to describe reasons \citep[on the latter, see][]{Parfit:1984fk}.
}
To illustrate, agent neutral representations are sometimes found at the early stages of planning.
Imagine that you and some friends are tasked with preparing a holiday.  
You might first write down a plan of action without specifying who will act; the plan simply describes what is to be done.
The plan  will eventually be implemented by you and your friends
 but this is not written in plan itself  and so it is agent-neutral.
Of course the fact that this plan is your collective plan may be represented elsewhere; this fact may also be implicit in  the plan's being stapled to the door of your communal kitchen.
The agent-neutrality of a representation does not require that the agents are nowhere specified, only that they are not specified in the content of the representation.%
\footnote{
Strictly the following argument does not hinge on the agent-neutrality of representations.
It is sufficient for our purposes that there are reciprocal motor representations  of outcomes whose obtaining would normally involve action on the part of each of the reciprocating agents.
In principle such representations could have contents which specify other agents or multiple agents. 
We focus on agent-neutral representations to show that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
}

Two or more agents have \emph{reciprocal} motor representations  just if there is a single outcome and each agent has a motor representation of that outcome. 
It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
[***Rest of this paragraph belongs in the evidence section?
Should be clearer about transition:
(a) I am observing and we reciprocally represent an outcome of your action (mirroring).
(b) We are interacting and reciprocally represent an outcome of your action (mirroring in joint action).
(c) We are interacting and reciprocally represent an outcome to which our actions are distributively directed.%
] 
It is more controversial that reciprocal motor representation occurs in joint action, but there is some evidence for this claim too.%
\footnote{
See \citet{kourtis:2012_predictive}: `the partner’s expected action is simulated at the motor level, which probably facilitates effective performance of the joint action.'
\citet{kourtis:2010_favoritism} show that 
reciprocal motor representation is more likely to occur in joint action than is mere observation. 
See also \citet{Knoblich:2003nf}.
 }

We want to go a tiny step further and suggest that in joint action there are sometimes reciprocal, agent-neutral representations of outcomes whose obtaining would normally involve action on the part of each of the agents.
So when we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions (and likewise for you).
For example, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
In this case I may represent the movement of the object from A to B and not only the component movements.
I represent a collective outcome of our actions and not just outcomes to which each of our actions are individually directed.
%
%So it is not just that, when we act together, I represent the outcomes to which your actions are directed and the outcomes to which my actions are directed; I may also represent an outcome to which all of our actions are directed.

We shall use the term \emph{social} motor representation as an abbreviation for the reciprocal, agent-neutral motor representation of outcomes whose obtaining would normally require action on the part of each reciprocating agent.
(This is probably a bad term to pick; we are still trying to come up with a better one. 
Certainly we cannot assume in advance of argument that such reciprocal, agent-neutral motor representations are social in any interesting sense;
for now we are using the term `social' in a non-standard way as an abbreviation, one that reflects our aim.)

What follows is speculative philosophy: we take for granted that  social motor representation sometimes facilitates joint action and ask whether this conjecture bears on our question about shared agency.
As a first step, we shall ask how social motor representation might facilitate joint action.
Note that answering this question does not directly commit us to any view about shared agency.
After all, social motor representation might be an enabling condition for some instances of shared agency but completely unnecessary for explaining what shared agency is.

>>>>

\section{Shared intention}
On the leading, best developed accounts of shared agency in philosophy and psychology, 
shared agency is explained  in terms of a special kind of intention or structure of intention, knowledge and commitment often called a shared intention.
For two or more agents to exercise shared agency is for them to have a shared intention and for this shared intention to be appropriately related to their actions.
Such actions comprise a \emph{shared intentional activity} (\citealp[p.\ 142]{Bratman:1999fr}).%
%On this widely held view, the difference between our walking together and two strangers merely walking side-by-side is that one event is appropriately related to a shared intention whereas the other is not.

@@

To avoid having to take sides on what shared intention is, 
let us abstract the details
and focus on features that everyone should agree on.
Minimally, a shared intention stands to the actions comprising a shared intentional activity in roughly the way that an ordinary, individual intention stands to an ordinary, individual action.
Now an ordinary, individual intention represents an outcome, coordinates an agent's actions, and coordinates the agent's actions in such a way that, normally, this coordination would facilitate the occurrence of the represented outcome.
Similarly, 
on almost any account,
a shared intention involves there being a single outcome which each agent represents, where these representations coordinate the several agents' actions and coordinate them in such a way that, normally, the coordination would facilitate the occurrence of the represented outcome.

There is another feature of shared intention that should be uncontroversial.  
Shared intentions or their components feature in practical reasoning alongside ordinary, individual intentions,
 and there are normative requirements which apply to combinations of individual and shared intentions.
To illustrate, tonight there is a party and a ceremony.
It is impossible for anyone to attend both, and this is common knowledge among us.
We have a shared intention that we attend the ceremony together.
While having this shared intention, I also intend to go to the party. 
Give our common knowledge,
this combination of shared and individual intentions is irrational. 
Its irrationality is related to that which would be involved in my individually intending to attend the ceremony while also intending to go to the party.
In short, shared intentions or their components are inferentially and normatively integrated with ordinary, individual intentions.

So much for shared intention.
Can we give a full account of shared agency by saying (roughly) that exercises of shared agency involve agents acting on a shared intention?
Or are further ingredients necessary to fully explicate shared agency?
Our view is that further ingredients are necessary.
In giving a full account of shared agency, it is also necessary to appeal to certain structures of motor representation. 
Note, however, that our view is consistent with various weaker claims linking shared intention to shared agency.
Our view is consistent with the claim that appeal shared intention is also needed in giving an account of shared agency, 
and with the claim that \emph{some forms} of shared agency can be fully characterised by appeal to shared intention alone.
Our view is also consistent with the claim that all exercises of shared agency involve shared intentions in some way.%
\footnote{
See \citet[p.\ 5]{Gilbert:2006wr}: `I take a collective action to involve a collective intention.'  
Related claims are made by 
	\citet[p.\ 381]{Carpenter:2009wq}, 
	\citet[p.\ 369]{Call:2009fk}, 
	\citet{Kutz:2000si},  
	\citet[p.\ 117]{rakoczy_pretend_2006},  and
	\citet{Tollefsen:2005vh}.
}

@@

First, if someone thought that it is possible to distinguish systematically between the contrast cases by appeal to shared intention (and perhaps only by appeal to shared intention), this could motivate holding that further ingredients are unnecessary in explicating shared agency.
We shall argue, however, that distinguishing systematically between the contrast cases cannot be done entirely in terms of shared intention. 
[*CV: this argument is missing from this draft.]

{\label{text:second_consideration} A second, related consideration} which might appear to favour this view is a parallel between shared agency and ordinary, individual agency.
If a full account of individual agency could be given by appeal to intention and other propositional attitudes with no appeal to motor representation,
then it would seem unlikely that giving a full account of shared agency must somehow involve appeal to motor representation.
However, as mentioned above, some have argued that fully characterising individual agency requires appeal to motor representation as well as intention \citep{pacherie:2000_content,butterfill:2012_intention}. 
Given these arguments, the parallel between individual and shared agency motivates considering the possibility that  there is a constitutive role for motor representation in shared agency. 
We also note that some philosophers allow that there could be agents whose actions are purposive---and, in some cases, even intentional---although  the agents have no intentions at all.
For instance, Bratman describes a creature who `acts on the basis of its beliefs and considered desires' only (not intentions) as doing things `intentionally' (\citeyear[p.\ 251]{bratman:2000_valuing}; see \citealp[pp.\ 137--8]{Bratman:1987xw} for further justification). 
So there are several grounds for doubting that a full account of ordinary, individual agency could be given just in terms of intention. 
If, as we believe, these doubts are well founded, then it is reasonable to ask whether ingredients other than shared intention are need to give a full account of shared agency.

\section{Motor representation}
Since we aim to argue that a full account of shared agency must appeal to motor representation, it may be helpful first to provide some background about motor representation.

@@

So much for the bare notion of motor representation.
Why suppose that motor representation might be needed in giving an account of shared agency? 
Our answer involves several steps. 
We first describe a structure of motor representation,
and then present evidence that such structures sometimes occur when two or more agents coordinate their actions. 
This will allow us to conclude only

\section{Social motor representation}
We start from an empirical premise about enabling conditions for shared agency: some joint actions are facilitated by reciprocal, agent-neutral motor representations of outcomes whose obtaining would normally involve action on the part of each agent.

A representation (motor or not) is \emph{agent-neutral} if its content does not specify any particular agent or agents.%
\footnote{
Our use of the term `agent-netural' to describe motor representations bears no relation to the use of the same term  to describe reasons \citep[on the latter, see][]{Parfit:1984fk}.
}
To illustrate, agent neutral representations are sometimes found at the early stages of planning.
Imagine that you and some friends are tasked with preparing a holiday.  
You might first write down a plan of action without specifying who will act; the plan simply describes what is to be done.
The plan  will eventually be implemented by you and your friends
 but this is not written in plan itself  and so it is agent-neutral.
Of course the fact that this plan is your collective plan may be represented elsewhere; this fact may also be implicit in  the plan's being stapled to the door of your communal kitchen.
The agent-neutrality of a representation does not require that the agents are nowhere specified, only that they are not specified in the content of the representation.%
\footnote{
Strictly the following argument does not hinge on the agent-neutrality of representations.
It is sufficient for our purposes that there are reciprocal motor representations  of outcomes whose obtaining would normally involve action on the part of each of the reciprocating agents.
In principle such representations could have contents which specify particular agents. 
However, we focus on agent-neutral representations to show that our view is consistent with the possibility that reciprocal motor representations are agent-neutral. 
}

Two or more agents have \emph{reciprocal} motor representations  just if there is a single outcome and each agent has a motor representation of that outcome. 
Reciprocal motor representations can occur both when one agent observes another acting alone and also when two or more agents act together.
Where on agent observes another acting alone,
it is possible that both have motor representations of an outcome to which the agent's action is directed.
For the agent, the representation plays a role in planning and monitoring; for the observer, the representation facilitates recognition.
This case of reciprocal motor representation---often called mirroring---has been extensively studied.
But our main interest is in a different case,
one where two or more agents are acting together,
 there is a single outcome to which their actions are directed,
 and the agents each have a motor representation of this outcome.
So in this case of reciprocal motor representation, the agents each represent an outcome whose obtaining would normally involve not only their own actions but also those of other agents.
(Note that our having reciprocal motor representations does not imply that we know this.)
When we act together, some of my motor representations may concern outcomes that are partly but not entirely to be realised by my actions; and likewise for you.
To illustrate, suppose our task is to move an object from A to B, where you pick it up and pass it to me so that I can then place it.
In this case I may represent the movement of the object from A to B and not only the component movements.
I represent a collective outcome of our actions and not just outcomes to which each of our actions are individually directed.
%
%So it is not just that, when we act together, I represent the outcomes to which your actions are directed and the outcomes to which my actions are directed; I may also represent an outcome to which all of our actions are directed.

We shall use the term \emph{shared} motor representation as an abbreviation for the reciprocal, agent-neutral motor representation of outcomes whose obtaining would normally require action on the part of each reciprocating agent. 
Note that shared motor representations are shared only in the sense in which two people can share a name.  
(We do not mean to suggest that they are shared in the sense in which two people can share a parent: there need be no representation with two or more subjects.)
Note also that a shared motor representation is not a special kind of motor representation: it is merely a structure of ordinary motor representations. 
At this stage we are not strictly entitled use the term `shared motor representation' as an abbreviation.
After all we have yet to show that these structures of motor representation play any constitutive role (or even any enabling role) in explaining shared agency.
The choice of term merely reflects our aim.

@@

It is hardly controversial that reciprocal motor representations exist, for their existence is suggested by by a large body of research on motor cognition in action observation.
[***Rest of this paragraph belongs in the evidence section?
Should be clearer about transition:
(a) I am observing and we reciprocally represent an outcome of your action (mirroring).
(b) We are interacting and reciprocally represent an outcome of your action (mirroring in joint action).
(c) We are interacting and reciprocally represent an outcome to which our actions are distributively directed.%
] 
It is more controversial that reciprocal motor representation occurs in joint action, but there is some evidence for this claim too.%
\footnote{
See \citet{kourtis:2012_predictive}: `the partner’s expected action is simulated at the motor level, which probably facilitates effective performance of the joint action.'
\citet{kourtis:2010_favoritism} show that 
reciprocal motor representation is more likely to occur in joint action than is mere observation. 
See also \citet{Knoblich:2003nf}.
 }

What follows is speculative philosophy: we now take for granted that  social motor representation sometimes facilitates the exercise of shared agency and ask whether this matters for a philosophical account of shared agency. 
As a first step, we shall ask how social motor representation might facilitate agents in coordinating their actions.
Note that answering this question does not directly commit us to any view about shared agency.
After all, social motor representation might sometimes facilitate coordination while being extraneous to a philosophical account of shared agency.
